---
title: "formatting exercise"
author: "Jingning Yang"
date: "9/17/2019"
output: pdf_document
 

---
## Extract from:  
Bradley Efron and Trevor Hastie   
*Computer Age Statistical Inference: Algorithms, Evidence, and Data Science      *
*Cambridge University Press, 2016   *
*<https://web.stanford.edu/~hastie/CASI_files/PDF/casi.pdf>*


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load("knitr")
```



Bradley Efron and Trevor Hastie
Computer Age Statistical Inference: Algorithms, Evidence, and Data Science
Cambridge University Press, 2016
<https://web.stanford.edu/~hastie/CASI_files/PDF/casi.pdf>

----------

<p style="text-indent:32px;"> Modern Bayesian practice uses various strategies to construct an appropriate “prior” $g\left ( \mu  \right )$in the absence of prior experience, leaving many statisticians unconvinced by the resulting Bayesian inferences. Our second example illustrates the difficulty.   

## Table 3.1 Scores from two tests taken by 22 students, mechanics and vectors.

```{r echo=FALSE}
library(xtable)
options(xtable.floating = FALSE)
options(xtable.timestamp = "")

chart = data.frame(mechanics=c(7,44,49,59,34,46,0,32,49,52,44),
                   vectors = c(51,69,41,70,42,40,40,45,57,64,61))
row.names(chart) <- (1:11)
kable(t(chart))   #t(chart)可以把竖着的表弄横。

chart1 = data.frame(mechanics=c(36,42,5,22,18,41,48,31,42,46,63),
                    vectors=c(59,60,30,58,51,63,38,42,69,49,63))
row.names(chart1) <- (12:22) 
kable(t(chart1))  


```





Table 3.1 shows the scores on two tests, mechanics and vectors, achieved by n D 22 students. The sample correlation coefficient between the two scores is $\hat{\theta }=0.498$,  

$$\hat{\theta} = \sum_{i=1}^{22}\left ( m_{i}-\bar{m}\right ) \left ( v_{i}-\bar{v} \right )\div \left [ \sum_{i=1}^{22}\left ( m_{i}-\bar{m} \right )^{2} \sum_{i=1}^{22}\left ( v_{i}-\bar{v} \right )^{2} \right ]^{1/2}$$
with $m$ and $v$ short for mechanics and vectors, $\bar{m}$ and $\bar{v}$ their averages.
